---
title: "Inferència estadística"
author: "Marc Comas i Marina Vives"
date: "15 de desembre"
output: ioslides_presentation
---

# Paràmetres de la població

## Variable numèrica

Què són els paràmetres $\mu$ i $\sigma$ d'una població $P$?

```{r, echo=FALSE, fig.height=2.5}
set.seed(1)
n = 1000
X = rnorm(n, mean=2)

p <- par(mar=c(2,2,1,1))
plot(X, cex=0.2, ylim=c(-2, 5), xlab = '')
legend('bottomleft', legend = c('Població'), col='black', lty =1, bty = 'n')
par(p)
```


## Variable numèrica

Què són els paràmetres $\mu$ i $\sigma$ d'una població $P$?

```{r, echo=FALSE, fig.height=2.5}
set.seed(1)
n = 1000
X = rnorm(n, mean=2)

p <- par(mar=c(2,2,1,1))
plot(X, cex=0.2, ylim=c(-2, 5), xlab = '')
legend('bottomleft', legend = c('Població'), col='black', lty =1, bty = 'n')
par(p)
```

\[
\begin{array}{rcl}
\mu &=& \frac{\sum_{x_i \in P} x_i }{ |P| } =  `r sprintf('%.2f', mean(X))` \\
\sigma^2 &=& \frac{\sum_{x_i \in P} (x_i-\mu)^2 }{ |P| } =  `r sprintf('%.2f', var(X))`
\end{array}
\]

## Variable numèrica

Què són els paràmetres $\mu$ i $\sigma$ d'una població $P$?

```{r, echo=FALSE, fig.height=2.5}
M = mean(X)
p <- par(mar=c(2,2,1,1))
plot(X, cex=0.2, ylim=c(-2, 5), xlab = '')
legend('bottomleft', legend = c('Població'), col='black', lty =1, bty = 'n')
segments(x0 = 1:n, x1 = 1:n, y0 = M, y = X, col='blue')
abline(h=M, col='red', lwd=3)
par(p)
```

\[
\begin{array}{rcl}
\mu &=& \frac{\sum_{x_i \in P} x_i }{ |P| } =  `r sprintf('%.2f', mean(X))` \\
\sigma^2 &=& \frac{\sum_{x_i \in P} (x_i-\mu)^2 }{ |P| } =  `r sprintf('%.2f', var(X))`
\end{array}
\]


## Variable numèrica

Què són els paràmetres de la població $\mu$ i $\sigma$?

```{r, echo=FALSE}
M = 3
MSError = sum((X-M)^2) / (n-1)
plot(X, cex=0.2, main=sprintf('Estimador: %.2f,  error mitjà %.3f', M, MSError))
abline(h=M, col='red')
segments(x0 = 1:n, x1 = 1:n, y0 = M, y = X, col='blue')
```
